---
title: "introduction to tidyR and Dplyr"
author: "Bram Sercu"
date: "October 19, 2017"
output: html_document
---


# 0. load packages and data

```{r}
library(dplyr)
library(tidyr)


```


# 1. create tibble and explore differences between tibble and dataframe
```{r}
# change dataframe to tible
cars_tib <- as_tibble(cars)

# type cars and cars_tib in the console
# show first 15 rows from cars_tib with print(cars_tib, n=12, width= Inf)

# create a similar dataframe and tibble and check the differences (copy to console)

tib1 <- tibble(
  numbers = c(1,2,3,4), 
  dutch = c("één", "twee", "drie", "vier"),
  english = c("one", "two", "three", "four")
)

df1 <- data.frame(
  numbers = c(1,2,3,4), 
  dutch = c("één", "twee", "drie", "vier"),
  english = c("one", "two", "three", "four")
)


tib1$numbers
df1$numbers

tib1$n
df1$n

tib1[, "numbers"]
df1[, "numbers"]

tib1[, c("numbers", "english")]
df1[, c("numbers", "english")]


```

# 2. tidy your data
use the command 'gather' when some of the column names are not names of variables, but values of a variable, to bring information that is spread over several columns to one column.
- http://r4ds.had.co.nz/tidy-data.html#gathering

Spreading is the opposite of gathering. You use it when an observation is scattered across multiple rows. 
- http://r4ds.had.co.nz/tidy-data.html#spreading

the functions 'seperate' and 'unite' let you, resprectively, split the content of a column in two columns or join columns according to a specified seperator.

## 2.1
```{r}
# load the dataset where the number of seedlings are counted at several dates (given in Julian Date)
# column 6 to 18 are the number of seedlings counted at that particular day (represented as julian day).

nr_seedlings <- read.table("data/nr_seedlings.txt", header=T)


nr_seedlings_tidy <- gather(nr_seedlings, key="julian_day", value = "nr_seedlings", JD233:JD399) 
nr_seedlings_tidy <- nr_seedlings %>% gather(key="julian_day", value = "nr_seedlings", JD233:JD399) 
```


## 2.2
```{r}
# dataset with lenght of inflorescence shoot, root and length/width of leaf
# create tidy dataframe with al measurements separate in a column.

plant_size <- read.table("data/plant_size.txt", header=T)

plant_size_tidy <- plant_size %>% separate()


```


## 2.3
```{r}
# unite enables to unite two columns two one
# reverse what you did in the previous exercise and combine the length and width of the leaf again to  one column using '_' as separator between the values

plant_size_less_tidy <- plant_size_tidy %<% unite()

```



# 3. work with you data
## 3.1 subsetting

We will work with the dataset 'nr_seedlings_tidy'
1. extract measurements that were taken before Julian day 331 
2. extract measurements for flowers 'b'
3. extract measurements that were taken before Julian day 331 for flowers 'b'
4. extract measurements that were taken before Julian day 331 for flowers 'b' and keep all columns except the column 'table'

```{r}
nr_seedlings_tidy_3.1.1 <- nr_seedlings_tidy %>% filter(julian_day < "JD331")
```

## 3.2 calculate new variables
Exercise for dataset 'plant_size_tidy'
1. calculate the leaf area
2. calculate the fraction of leaf length to inflorescence length
3. calculate normalized values for each variable.

```{r}

```

## 3.3 summarise
Exercise for dataset 'plant_size_tidy'
1. calculate average length of roots
2. calculate average length of all columns
3. calculate average length of roots per plot
4. calculate average length of all columns per plot

```{r}

```


## 3.4 combinations

Exercise for dataset 'nr_seedlings_tidy'
1. calculate total nr of seedlings per plant per julian day as new base dataset
2. calculate total nr of seedlings per inflorescence from the new dataset
3. calculate total nr of seedlings per plant
4. calculate the cumulative nr seedlings per plant from the lowest julian day to the highest (inspect the results!)
5. remove rows that contains NA as value for nr_seedlings (and run nr 4 again)

6. Join the datatable 'nr_seedlings_tidy' with 'plant_size_tidy' and keep only plants that are available in 'plant_size_tidy'
7. as nr 6 but keep all plants from both datatables.

```{r}
nr_seedlings_tidy3.4.1 <-


```



