---
title: "ggplot-excersises"
output: html_document
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## R ggplot2 Graphics: Exercises

#Create plots using data set "Salaries" from "car" package

#1. Plot the salary as a function of the number of years since obtaining a PhD
  - Group per sex (males: black, females:red) 
  - Add x-axis title: "Years since PhD"
  - Add y-axis title: "Salary"
  - Add main title: "Salaries for professors"
  - Add legend above the graph 
  
```{r}

```
  

#2. Make a boxplot with salaries as are shown for the different sexes per rank


```{r}

```



#3. Make the same graph as in #2, change theme
- Change line colours of the boxes
- Change the fill colours of the boxes
- Make the y-axis title bold
- Remove the x-axis tile

```{r}

```

#4. Combine the three previous plots into a multipanel plot, with the 3 graphs next to each other
- add a general title

```{r}
library(gridExtra)

```




#5. Make the first graph using qplot
```{r}

```
