---
title: "ggplot-solutions"
output: html_document
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## R ggplot2 Graphics: Exercises

#Create plots using data set "Salaries" from "car" package

#1. Plot the salary as a function of the number of years since obtaining a PhD
  - Group per sex (males: black, females:red) 
  - Add x-axis title: "Years since PhD"
  - Add y-axis title: "Salary"
  - Add main title: "Salaries for professors"
  - Add legend above the graph 
  
```{r}
ggplot(Salaries,aes(x=yrs.since.phd,y=salary,colour=sex)) +
  geom_point() +
  labs(title="Salaries for professors",x="Years since PhD",y="Salary") +
  theme(legend.position="top")
```
  

#2. Make a boxplot with salaries as are shown for the different sexes per rank


```{r}
ggplot(Salaries,aes(x=rank,y=salary,colour=sex)) +
  geom_boxplot() +
  labs(title="Salaries for professors",x="Rank",y="Salary")
```

Alternative
```{r}
ggplot(Salaries,aes(x=sex,y=salary,colour=sex)) +
  geom_boxplot() +
  labs(title="Salaries for professors", x="Sex",y="Salary") +
  facet_grid(.~rank) + 
  theme(legend.position="none")
```


#3. Make the same graph as in #2, change theme
- Change line colours of the boxes
- Change the fill colours of the boxes
- Make the y-axis title bold
- Remove the x-axis tile

```{r}
ggplot(Salaries,aes(x=sex,y=salary,fill=sex)) +
  geom_boxplot() +
  labs(title="Salaries for professors", x="Sex",y="Salary") +
  facet_grid(.~rank) + 
  theme(legend.position="none",
        axis.title.y = element_text(face="bold"),
        axis.title.x = element_blank())
```

#4. Combine the three previous plots into a multipanel plot, with the 3 graphs next to each other
- add a general title
```{r}
library(gridExtra)

one = ggplot(Salaries,aes(x=yrs.since.phd,y=salary,colour=sex)) +
  geom_point() +
  labs(x="Years since PhD",y="Salary") +
  theme(legend.position="top")

two = ggplot(Salaries,aes(x=sex,y=salary,colour=sex)) +
  geom_boxplot() +
  labs(x="Sex",y="Salary") +
  facet_grid(.~rank) + 
  theme(legend.position="none")

three = ggplot(Salaries,aes(x=sex,y=salary,fill=sex)) +
  geom_boxplot() +
  labs(x="Sex",y="Salary") +
  facet_grid(.~rank) + 
  theme(legend.position="none",
        axis.title.y = element_text(face="bold"),
        axis.title.x = element_blank())

grid.arrange(one, two, three,
             ncol=3,
             top="Salaries for professors")


```




#5. Make the first graph using qplot
```{r}
qplot(Salaries,aes(x=yrs.since.phd,y=salary,colour=sex)) +
  geom_point() +
  labs(title="Salaries for professors",x="Years since PhD",y="Salary") +
  theme(legend.position="top")
```
