---
title: "introduction to tidyR and Dplyr"
author: "Bram Sercu"
date: "October 19, 2017"
output: html_document
---


# 0. load packages and data

```{r}
library(dplyr)
library(tidyr)
```


# 1. create tibble and explore differences between tibble and dataframe

```{r}
# change dataframe to tible
cars_tib <- as_tibble(cars)

# type cars and cars_tib in the console
# show first 15 rows from cars_tib with print(cars_tib, n=12, width= Inf)

# create a similar dataframe and tibble and check the differences (copy to console)

tib1 <- tibble(
  numbers = c(1,2,3,4), 
  dutch = c("één", "twee", "drie", "vier"),
  english = c("one", "two", "three", "four")
)

df1 <- data.frame(
  numbers = c(1,2,3,4), 
  dutch = c("één", "twee", "drie", "vier"),
  english = c("one", "two", "three", "four")
)


tib1$numbers
df1$numbers

tib1$n
df1$n

tib1[, "numbers"]
df1[, "numbers"]

tib1[, c("numbers", "english")]
df1[, c("numbers", "english")]


```

# 2. tidy your data

## 2.1, Using gather

use the command 'gather' when some of the column names are not names of variables, but values of a variable, to bring information that is spread over several columns to one column.
- http://r4ds.had.co.nz/tidy-data.html#gathering

```{r}
# load the dataset where the number of seedlings are counted at several dates (given in Julian Date)
# column 6 to 18 are the number of seedlings counted at that particular day (represented as julian day).

nr_seedlings <- read.table("data/nr_seedlings.txt", header=T)
```

Make this data tidy:

```{r}

### your code here:

```


## 2.2 Using spread

Spreading is the opposite of gathering. You use it when an observation is scattered across multiple rows. 
- http://r4ds.had.co.nz/tidy-data.html#spreading

```{r}
# dataset with lenght of inflorescence shoot, root and length/width of leaf
plant_size <- read.table("data/plant_size.txt", header=T)

```

Make this data tidy:

```{r}

### Your code here

```


## 2.3 Using separate and unite

The functions 'seperate' and 'unite' let you, resprectively, split the content of a column in two columns or join columns according to a specified seperator.

From the dataset in 2.2 separate the width/length measurement into two separate columns:

```{r}

### Your code here

```

Now re-unite the width / length measurements into a new column: "width_length" (separator "_"):

```{r}

### Your code here

```

# 3. dplyr

## 3.1 filter

In the "plant_size" dataset, filter rows to keep only observations from plot A

```{r}

### Your code here

```

## 3.2 arrange

From the tidy plant size dataset arrange plants by inflorescence length

```{r}

### Your code here

```

Tipp, here is how to get this tidy data

```{r}
plant_size_tidy <- spread(plant_size, part, length)
plant_size_tidy <- separate(plant_size_tidy, col = leaf, into = c("leaf_length","leaf_width"),sep="/")
```


## 3.3 select

From the tidy plant size only keep the Plant, Plot and two leaves columns

```{r}

### Your code here

```


## 3.4 mutate

From the tidy plant size compute the leaf area (assuming leaves are rectangles)

```{r}

### Your code here

```


## 3.5 summarise + group_by

From the tidy plant size compute the average and median root length per plot

```{r}

### Your code here

```




# 3. work with you data
## 3.1 subsetting

We will work with the dataset 'nr_seedlings_tidy'
1. extract measurements that were taken before Julian day 331 
2. extract measurements for flowers 'b'
3. extract measurements that were taken before Julian day 331 for flowers 'b'
4. extract measurements that were taken before Julian day 331 for flowers 'b' and keep all columns except the column 'table'

```{r}
nr_seedlings_tidy_3.1.1 <- nr_seedlings_tidy %>% filter(julian_day < "JD331")
```

## 3.2 calculate new variables
Exercise for dataset 'plant_size_tidy'
1. calculate the leaf area
2. calculate the fraction of leaf length to inflorescence length
3. calculate normalized values for each variable.

```{r}

```

## 3.3 summarise
Exercise for dataset 'plant_size_tidy'
1. calculate average length of roots
2. calculate average length of all columns
3. calculate average length of roots per plot
4. calculate average length of all columns per plot

```{r}

```


## 3.4 combinations

Exercise for dataset 'nr_seedlings_tidy'
1. calculate total nr of seedlings per plant per julian day as new base dataset
2. calculate total nr of seedlings per inflorescence from the new dataset
3. calculate total nr of seedlings per plant
4. calculate the cumulative (from the lowest julian day to the highest) nr seedlings per plant  (inspect the results!)
5. remove rows that contains NA as value for nr_seedlings (and run nr 4 again)
6. Join the datatable 'nr_seedlings_tidy_3.4.4' with 'plant_size_tidy' and keep only plants that are available in 'plant_size_tidy'
7. as nr 6 but keep all plants from both datatables.

```{r}
nr_seedlings_tidy3.4.1 <-


```

