---
title: "Back into the tidyverse"
author: "Lionel Hertzog & Bram Sercu"
date: "September 19, 2018"
output: html_document
---


# 0. load packages and data

```{r}
library(dplyr)
library(tidyr)
```

# 1. tidy your data

## 1.1, Using gather

use the command 'gather' when some of the column names are not names of variables, but values of a variable, to bring information that is spread over several columns to one column.
- http://r4ds.had.co.nz/tidy-data.html#gathering

```{r}
# load the dataset where the number of seedlings are counted at several dates (given in Julian Date)
# column 6 to 18 are the number of seedlings counted at that particular day (represented as julian day).

nr_seedlings <- read.table("data/nr_seedlings.txt", header=T)
```

Make this data tidy:

```{r}

### your code here:

```


## 1.2 Using spread

Spreading is the opposite of gathering. You use it when an observation is scattered across multiple rows. 
- http://r4ds.had.co.nz/tidy-data.html#spreading

```{r}
# dataset with lenght of inflorescence shoot, root and length/width of leaf
plant_size <- read.table("data/plant_size.txt", header=T)

```

Make this data (un)tidy:

```{r}

### Your code here

```


## 1.3 Using separate and unite

The functions 'seperate' and 'unite' let you, resprectively, split the content of a column in two columns or join columns according to a specified seperator.

From the dataset in 1.2 separate the width/length in leaf measurement into two separate columns:

```{r}

### Your code here

```

Now re-unite the width / length measurements into a new column: "width_length" (separator "_"):

```{r}

### Your code here

```

# 2. dplyr

## 2.1 filter

In the "plant_size" dataset, filter rows to keep only observations from plot A

```{r}

### Your code here

```

## 2.2 arrange

From the tidy plant size dataset arrange plants by inflorescence length

```{r}

### Your code here

```

Tipp, here is how to get this tidy data

```{r}
plant_size_tidy <- spread(plant_size, part, length)
plant_size_tidy <- separate(plant_size_tidy, col = leaf, into = c("leaf_length","leaf_width"),sep="/")
```


## 2.3 select

From the tidy plant size only keep the Plant, Plot and two leaves columns

```{r}

### Your code here

```


## 2.4 mutate

From the tidy plant size compute the leaf area (assuming leaves are rectangles)

```{r}

### Your code here

```


## 2.5 summarise + group_by

From the tidy plant size compute the average and median root length per plot

```{r}

### Your code here

```




# 3. Piping - more advanced manipulation

## 3.1 

From nr_seedlings, do the following (using pipes if you feel like it):
1. tidy the dataset (gather ...)
2. Create a new variable: "julian_day" from the number in the column names, ie: from JD233 put 233 in this julian_day column, keep this dataset you will need it later on
3. Only keep columns: plant, inflorescence, plate, julian_day and value
4. Compute the total number of observations per plants

```{r}

### Your code here

```


## 3.2

From plant_size, do the following:
1. tidy the dataset (using spread), also separating leaf length from leaf width
2. compute leaf area
3. compute normalized values for all variables

```{r}

### Your code here

```

## 3.3

From the dataset obtained in 3.1, do the following:
1. calculate total number of seedlings per plant and per julian day
2. arrange per plant and per julian day 
3. calculate the cumulative number of seedlings per plant and per julian day
4. only keep plant id 1:10

```{r}

### Your code here

```


## Bonus I

From the nr_seedling dataset

1. based on the results of 3.3 compute the julian_day when the maximum number of flower was reached
2. joined these results with the plant_size table
3. explore relationships between plant sizes and the time of maximum flowering (scatterplots ...)

```{r}

## Your code here

```

## Bonus II: create tibble and explore differences between tibble and dataframe

```{r}
# change dataframe to tible
cars_tib <- as_tibble(cars)

# type cars and cars_tib in the console
# show first 15 rows from cars_tib with print(cars_tib, n=12, width= Inf)

# create a similar dataframe and tibble and check the differences (copy to console)

tib1 <- tibble(
  numbers = c(1,2,3,4), 
  dutch = c("één", "twee", "drie", "vier"),
  english = c("one", "two", "three", "four")
)

df1 <- data.frame(
  numbers = c(1,2,3,4), 
  dutch = c("één", "twee", "drie", "vier"),
  english = c("one", "two", "three", "four")
)


tib1$numbers
df1$numbers

tib1$n
df1$n

tib1[, "numbers"]
df1[, "numbers"]

tib1[, c("numbers", "english")]
df1[, c("numbers", "english")]


```


