---
title: "Simulation coding club"
author: "Lionel Hertzog"
date: "May 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation B, A, BA

Simulation as we under

The r* functions

Simple examples with rbinom or rnorm (sex of offsprings or height), show uncertainties in simulation process, importance of replications

Some technicalities about simulation in R, looping, vectorization, replicate

### Exercice


## Using simulations for testing model fitness

Based on blog posts and DHARMa

### Exercice

## Using simulations for power analysis

An easy and more tricky exmaple inspired from Bolker's book

### Exercice

```{r}
n <- seq(10,100,10)
pow_fun <- function(n){
  x <- runif(n,-2,2)
  y <- rpois(n,lambda=exp(0.5 + 0.25 * x))
  m <- glm(y~x,family = "poisson")
  return(summary(m)$coefficients[2,4])
}
n_rep <- 100
pow_slp <- sapply(n,function(n) sum(replicate(n_rep,pow_fun(n)) < 0.05) / n_rep)
{plot(n,pow_slp,type="l")
abline(h=0.8,lty=2,lwd=2)}
```

## Using simulations to explore processes

Some movement or other examples from the working group

### Exercice










